Composable Concurrency Models

### 2016-11-14 slides/slides.tex
Wrote background slides

\usepackage{graphicx}
\usepackage{tikz}

@@

@@

        \onslide<3>{
          \node at (0,-0.62) {
            \includegraphics[width=220pt]{goals}
          };
        }
      \end{tikzpicture}
    \end{figure}
  \end{frame}
\end{document}

>>>>

\usepackage[USenglish]{babel}
\usepackage[backend=bibtex,citestyle=authortitle]{biblatex}
\usepackage{graphicx}
\usepackage{tikz}

@@

\bibliography{references}

@@

        \onslide<3>{
          \node at (0,-0.62) {
            \includegraphics[width=220pt]{goals}
          };
        }
      \end{tikzpicture}
    \end{figure}
  \end{frame}

  \section{Background}

  \begin{frame}
    \frametitle{Concurrency}
    \framesubtitle{Intuitive definition}

    \onslide<1>{
      \begin{columns}[T]
        \begin{column}{0.48\textwidth}
          \centering
          sequential
        \end{column}

        \hfill

        \begin{column}{0.48\textwidth}
          \centering
          concurrent
        \end{column}
      \end{columns}
    }

    \onslide<2>{
      \begin{columns}[T]
        \begin{column}{0.48\textwidth}
          \centering
          concurrent
        \end{column}

        \hfill

        \begin{column}{0.48\textwidth}
          \centering
          parallel
        \end{column}
      \end{columns}
    }
  \end{frame}

  \begin{frame}
    \frametitle{Concurrency}
    \framesubtitle{Formal definition}

    \textbf{The ``happens before'' ($\rightarrow$) relation~\footcite{Lamport1977}}

    $A \rightarrow B$ if one of the following is true:

    \begin{enumerate}
      \item $A$ and $B$ are operations in the same thread and $A$ occurs before $B$.
      \item $A$ is the sending of a message by one thread and $B$ is the receipt of the same message by another thread.
    \end{enumerate}

    $A$ and $B$ are said to be concurrent if $A \nrightarrow B$ and $B \nrightarrow A$.
  \end{frame}

  \begin{frame}
    \frametitle{Consistency models}

    \begin{itemize}
      \item \textbf{Sequential:} Does the history of the program yield a correct result?
      \item \textbf{Concurrent:} Does every possible history of the program yield a correct result?
    \end{itemize}
  \end{frame}

  \begin{frame}
    \frametitle{Complications}

    \onslide<1->{
      \centering
      \includegraphics{single-thread-read-write}
    }

    \onslide<2->{
      \centering
      \includegraphics{multiple-threads-read-write}
    }
  \end{frame}

  \begin{frame}
    \frametitle{Complications}

    \begin{figure}
      \centering
      \includegraphics{noninstantaneity}
    \end{figure}

    \begin{itemize}
      \item Nothing is truly instantaneous.
      \item The history of a program may be influenced by the time it takes for operations to complete.
    \end{itemize}
  \end{frame}

  \begin{frame}
    \frametitle{Linearizability}

    \begin{figure}
      \centering
      \includegraphics{linearizability}
    \end{figure}

    \begin{itemize}
      \item Linearizability guarantees that the completion of an operation on a single object will appear to be instantaneous.
      \item The results of a linearizable operation will be visible as soon as the operation is complete.
    \end{itemize}
  \end{frame}

  \begin{frame}
    \frametitle{Serializability}

    \begin{figure}
      \centering
      \includegraphics{serializability}
    \end{figure}

    \begin{itemize}
      \item Serializability guarantees that operations can occur in any order as long as an equivalent serial history exists.
      \item While a serializable set of operations is being executed, it appears to be the only set of operations being executed.
    \end{itemize}
  \end{frame}

  \begin{frame}
    \frametitle{Strict serializability}

    \begin{itemize}
      \item Linearizability is a guarantee of consistency.
      \item Serializability is a guarantee of isolation.
      \item Together, linearizability and serializability guarantee strict serializability---a history is equivalent to some serial execution order and that order corresponds to the execution order in real time.
    \end{itemize}
  \end{frame}

\end{document}
